---
title: "p8105_hw3_rz2570"
author: "Ruilian Zhang"
date: "10/12/2021"
output: github_document
---

```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Problem 1

```{r}
library(p8105.datasets)
data("instacart")

dim(instacart)

names(instacart)

head(instacart) 
```

* The `instacart` dataframe has `r ncol(instacart)` of columns and `r nrow(instacart)` of observations. Each row contains relative information of an item from one order , which are: `r names(instacart)`.  
* For example, for product `49302`, its order of being added to cart was `1` by user `112108`, and it's from `dairy eggs` department.  

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    n_obs = n()) %>% 
  arrange(-n_obs) %>% 
  nrow()
```

* There are 134 aisles, the `fresh vegetables` (`aisle_id` is 83) aisle is the most items are ordered from.

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    n_obs = n()
  ) %>% 
  filter(n_obs > 10000) %>% 
  ggplot(aes(x = aisle, y = n_obs)) +
  geom_bar(stat = "identity")
```

